Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-04-30 RNARedPrint-Bioinfo2018.tex
New Figure 2A and corresponding updates to the caption and text

Fig.~\ref{fig:energydist} illustrates the described
strategy. \todo{Update Figure 2A}Fig.~\ref{fig:energydist}A
illustrates our capability to target Turner energies based on the
stacking energy model. It shows the relative distance between
$\EnergyTurner$ and $\EnergyStacking$, observed for instances of the
\texttt{Modena} benchmark. For each structure, $n=1\,000$ sequences
were sampled, while targeting a ${\sf GC}\%$ of $50\%$, and reported
the difference between the expected and sampled energies for both the
Turner model and its approximation using a stacking model.  For an
example instance of the \texttt{Modena} benchmark \texttt{3str},
Fig.~\ref{fig:energydist}B shows the Turner energy distributions of
the single structures as they result from sampling with different weight
parameters. The figure illustrates how our multidimensionnal Boltzmann
sampling strategy can accurately and, to a large extent, independently
shift the Turner energies of sampled sequences to prescribed targets.

%To reduce the cost of the precomputation, we used a stacking energy model 
%as a proxy for the Turner model. Indeed, we observed a strong linear 
%dependency between the Turner energy and the 
%emphasize that only the simple stacking energies are directly targeted
%(at $\pm$10\% tolerance, except for a few hard instances). Due to our
%linear adjustment, we finally achieve well-controlled energies in the
%realistic Turner energy model.
\begin{figure*}[t]
  \begin{center}
    {\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset}\hspace{.5cm}
    {\sf \bfseries B}\includegraphicstop[width=0.57\textwidth]{Figs/energy_distribution}
  \end{center}
  \caption{%
    Targeting specific energies using multi-dimensional Boltzmann
    sampling.  (A) Accuracy of targeting over all target structures of
    the \Software{Modena} benchmark instances \texttt{RNAtabupath}
    (2str), \texttt{3str} and \texttt{4str}. Shown is the relative
    deviation of targeted and achieved energies in the simple stacking
    model and the Turner model.  (B) Turner energy distributions for
    three target structures S1, S2 and S3. We show the results of
    targeting the respective free energies $(-40,-40,-20)$ and
    $(-20,-20,-20)$ for the target structures (annotated in red and
    blue, resepectively)---this demonstrates the
    effectivity of our adaptive multi-dimensional Boltzmann sampling
    procedure; moreover, for comparison, the distributions for uniform
    (U) and Boltzmann (B) samples---respectively associated with
    homogenous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
      }
  \label{fig:energydist}
\end{figure*}
\subsection{Generating high-quality seeds for further optimization}
We empirically study the capacity of \ourprog{} to improve the generation of seed sequences for subsequent local optimizations, showing an important application in biologically relevant scenarii.

@@

Using our strategy, we generated at least 1\,000 seeds per instance of
the subsets \texttt{RNAtabupath} (2str), \texttt{3str}, and \texttt{4str}, of
the \Software{Modena} benchmark set with 2,3, and 4 structures. The
results are compared, in terms of their value for the objective
function of Eq.~\eqref{eq:blueprintobjective}, to seed sequences
uniformly sampled using \RNAblueprint{}~\citep{Hammer2017}
(Table~\ref{tab:benchmark-results}, seeds). Our first analysis reveals
that Boltzmann-sampled sequences constitute better seeds, associated
with better values ($\sim\! 69\%$ improvement on average), compared to
uniform seeds for \emph{all instances} of our benchmark
(Supp. Mat.~\ref{sec:validity}).

>>>>

Fig.~\ref{fig:energydist} illustrates the described
strategy. Fig.~\ref{fig:energydist}A
illustrates our capability to target Turner energies based on the
stacking energy model. It shows the deviation between
$\EnergyTurner$ and $\EnergyStacking$, observed for instances of the
\texttt{Modena} benchmark. For each structure, $n=1\,000$ sequences
were sampled, while targeting a ${\sf GC}\%$ of $50\%$, and reported
the difference between the expected and sampled energies for both the
Turner model and its approximation using a stacking model.  For an
example instance of the \texttt{Modena} benchmark \texttt{3str},
Fig.~\ref{fig:energydist}B shows the Turner energy distributions of
the single structures as they result from sampling with different weight
parameters. The figure illustrates how our multidimensionnal Boltzmann
sampling strategy can accurately and, to a large extent, independently
shift the Turner energies of sampled sequences to prescribed targets.

%To reduce the cost of the precomputation, we used a stacking energy model 
%as a proxy for the Turner model. Indeed, we observed a strong linear 
%dependency between the Turner energy and the 
%emphasize that only the simple stacking energies are directly targeted
%(at $\pm$10\% tolerance, except for a few hard instances). Due to our
%linear adjustment, we finally achieve well-controlled energies in the
%realistic Turner energy model.
\begin{figure*}[t]
  \begin{center}
    {\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset2}\hspace{.5cm}
    {\sf \bfseries B}\includegraphicstop[width=0.57\textwidth]{Figs/energy_distribution}
  \end{center}
  \caption{%
    Targeting specific energies using multi-dimensional Boltzmann
    sampling.  (A) Accuracy of targeting over all target structures of
    the \Software{Modena} benchmark instances \texttt{RNAtabupath}
    (2str), \texttt{3str} and \texttt{4str}. 
    For the stacking and the Turner energy model, the relative offsets between the targeted and 
    achieved energies were determined. We show the deviations of these offsets between both models.
    (B) Turner energy distributions for
    three target structures S1, S2 and S3. We show the results of
    targeting the respective free energies $(-40,-40,-20)$ and
    $(-20,-20,-20)$ for the target structures (annotated in red and
    blue, resepectively)---this demonstrates the
    effectivity of our adaptive multi-dimensional Boltzmann sampling
    procedure; moreover, for comparison, the distributions for uniform
    (U) and Boltzmann (B) samples---respectively associated with
    homogenous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{fig:energydist}
\end{figure*}
\subsection{Generating high-quality seeds for further optimization}
We empirically study the capacity of \ourprog{} to improve the generation of seed sequences for subsequent local optimizations, showing an important application in biologically relevant scenarii.

@@

Using our strategy, we generated at least 1\,000 seeds per instance of
the subsets \texttt{RNAtabupath} (2str), \texttt{3str}, and
\texttt{4str}, of the \Software{Modena} benchmark set with 2,3, and 4
structures. The results are compared, in terms of their value for the
objective function of Eq.~\eqref{eq:blueprintobjective}, to seed
sequences uniformly sampled using \RNAblueprint{}~\citep{Hammer2017}
(Table~\ref{tab:benchmark-results}, seeds). Our first analysis reveals
that Boltzmann-sampled sequences constitute better seeds
($\sim\! 69\%$ improvement on average) compared to uniform seeds for
\emph{all instances} of our benchmark (Supp. Mat.~\ref{sec:validity}).
